Discrete Quantum Theories and Computing

### 2019-01-08 dissertation.tex
dissertation.lyx "Chapter 6 Further Discussion" Updating...

The quantum extension, quantum interval-valued probability measure
(QIVPM) $\bar{\mu}:\events\rightarrow\mathscr{I}$, is a generalization
of both classical IVPMs $\bar{\mu}:2^{\Omega}\rightarrow\mathscr{I}$
and conventional quantum probability measures $\mu:\events\rightarrow[0,1]$
\cite{THOS2017}, because QIVPMs reduce to classical IVPMs when the
space of quantum events $\events$ is restricted to mutually commuting
events, and reduces to conventional quantum probability measures when
mapping to infinitely precise uncountable intervals $\mathscr{I}_{\infty}=\set{\left[x,x\right]}{x\in\left[0,1\right]}$.
While Shapley and Gleason both proved there must be a ``state''
consistent with any given QIVPM in the reduced cases, in general there
exists a QIVPM such that no state is consistent with it. However,
we found a class of QIVPMs such that all QIVPMs in this class are
consistent with a non-empty ``ball'' of quantum states whose radius
is defined by the maximal length of the intervals, and recovers the
original Gleason theorem asymptotically. Similarly, the conventional
quantum expectation value and the classical Choquet integral are together
generalized to the quantum interval-valued expectation value. This
is used to prove an imprecise Kochen-Specker theorem \cite{BELL_1966,kochenspecker1967,Redhead1987-REDINA,peres1995quantum,Jaeger2007}
which suggests a possible resolution of the Meyer-Mermin debate on
the impact of finite-precision measurement on the Kochen-Specker theorem
\cite{PhysRevLett.83.3751,Mermin1999}.
\end{comment}
\begin{comment}
We present a mathematical framework based on quantum interval-valued
probability measures to study the effect of experimental imperfections
and finite precision measurements on defining aspects of quantum mechanics
such as contextuality and the Born rule. While foundational results
such as the Kochen-Specker and Gleason theorems are valid in the context
of infinite precision, they fail to hold in general in a world with
limited resources. Here we employ an interval-valued framework to
establish bounds on the validity of those theorems in realistic experimental
environments. In this way, not only can we quantify the idea of finite-precision
measurement within our theory, but we can also suggest a possible
resolution of the Meyer-Mermin debate on the impact of finite-precision
measurement on the Kochen-Specker theorem.
\end{comment}
Sec.~\ref{sec:Interval-Uncertainty} starts from reviewing how to
formulate classical ``imprecise probabilities'' \cite{Shafer1976,GilboaSchmeidler1994,Marinacci1999,JamisonLodwick2004,HuberRonchetti2009,Grabisch2016}
as interval-valued probability measures (IVPMs). Given a classical
IVPM, its core is the set of the real-valued probability measures
consistent with it \cite{Shapley1971,GilboaSchmeidler1994,NgMoYeh1997Chinese,Grabisch2016},
and the expectation value with respect to it is just classical Choquet
integral \cite{Choquet1954,GilboaSchmeidler1994,Grabisch2016}. These
classical properties can naturally be extended for QIVPMs when we
extend classical IVPMs to QIVPMs. In Sec.~\ref{sec:Kochen-Specker},
we utilize the previous defined expectation values while proving the
finite precision variants of the Kochen-Specker theorem \cite{BELL_1966,kochenspecker1967,Redhead1987-REDINA,peres1995quantum,Jaeger2007},
and establish bounds on its validity in realistic experimental environments.
In Sec.~\ref{sec:Gleason}, we are able to show that, while a QIVPM
incorporating the effects of finite precision might not be consistent
with Gleason's unique state~$\rho$ on all projectors defined on
a Hilbert space~$\Hilb$ of dimension $d\ge3$, there is a mathematically
precise sense in which one recovers the original Gleason theorem asymptotically.
Specifically, it is possible to construct a class of QIVPMs representing
bounded resources that is parameterized by the size of the intervals.

@@

\begin{tabularx}{\linewidth}{YYYYYYY}
\toprule 
\addlinespace
 & Conventional & Modal & Discrete (I) & Discrete (II) & QPMFF & QIVPM\tabularnewline\addlinespace
\addlinespace
\midrule
\midrule 
\addlinespace
 & Chapter~\ref{chap:Conventional-Quantum-Theory} & Sec.~\ref{modalquantum} & Sec.~\ref{discretequantumtheoryI} to \ref{sec:Geometry-of-Entangled-States} & Sec.~\ref{sec:Discrete-Quantum-Theory-(II)} & Sec.~\ref{sec:Toward-IVPM} & Chapter~\ref{chap:QIVPM}\tabularnewline\addlinespace
\addlinespace
\midrule
\addlinespace
States space & $\mathbb{C}^{D}$ & $\Fq^{D}$ & $\Fpp^{D}$ & Local region in $\Fpp^{D}$ & $\Fpp^{D}$ & $\mathbb{C}^{D}$\tabularnewline\addlinespace
\addlinespace
\midrule
\addlinespace
Likelihood of events is predicted by & Real-valued probability & Possible or impossible & Possible or impossible & Cardinal probability & Real-valued probability, but no sensible Born rule & Interval-valued probability\tabularnewline\addlinespace
\addlinespace
\midrule
\addlinespace
Expectation value & Defined & Undefined & Formally defined & Undefined &  & Defined\tabularnewline\addlinespace
\addlinespace
\midrule
\midrule 
\addlinespace
 &  & Sec.~\ref{modalquantumcomputing} & Sec.~\ref{discretequantumcomputingI} & Sec.~\ref{discretequantumcomputingII} &  & \tabularnewline\addlinespace
\addlinespace
\midrule
\addlinespace
Deutsch's algorithm & Yes & Maybe no & Yes & Yes &  & \tabularnewline\addlinespace
\addlinespace
\midrule
\addlinespace
Efficiently solve $\usat$ & Unlikely & Yes & Partially & Unlikely &  & \tabularnewline\addlinespace
\addlinespace
\midrule
\addlinespace
Grover search algorithm & Yes &  &  & Yes &  & \tabularnewline\addlinespace
\addlinespace
\bottomrule
\end{tabularx}
\end{sidewaystable}
The organization of this thesis is summarized in Table~\ref{tab:organization},
and we present our conclusions in final Chapter~\ref{chap:Conclusion}.

@@

@@

\todo{Add a two-page conclusion chapter which could include further
discussion, but it can be only a few sentences? }\todo{Discrete
Schrödinger equation? And its relation with Grover's algorithm? }\todo{Notice
that in differential geometry, Killing–Hopf theorem asserts that having
positive constant curvature guarantees a manifold is essentially a
sphere, but say little things even if the curvature has only small
deviation from constant. Since a quantum probability measure is glued
by classical probability measures, like a manifold. When their relations
are exact, Gleason's theorem asserts that a quantum probability measure
can essentially induced by the Born rule, but they can be wild when
we move on to interval-valued situation! }

>>>>

The quantum extension, quantum interval-valued probability measure
(QIVPM) $\bar{\mu}:\events\rightarrow\mathscr{I}$, is a generalization
of both classical IVPMs $\bar{\mu}:2^{\Omega}\rightarrow\mathscr{I}$
and conventional quantum probability measures $\mu:\events\rightarrow[0,1]$
\cite{THOS2017}, because QIVPMs reduce to classical IVPMs when the
space of quantum events $\events$ is restricted to mutually commuting
events, and reduces to conventional quantum probability measures when
mapping to infinitely precise uncountable intervals $\mathscr{I}_{\infty}=\set{\left[x,x\right]}{x\in\left[0,1\right]}$.
While Shapley and Gleason both proved there must be a ``state''
consistent with any given QIVPM in the reduced cases, in general there
exists a QIVPM such that no state is consistent with it. However,
we found a class of QIVPMs such that all QIVPMs in this class are
consistent with a non-empty ``ball'' of quantum states whose radius
is defined by the maximal length of the intervals, and recovers the
original Gleason theorem asymptotically. Similarly, the conventional
quantum expectation value and the classical Choquet integral are together
generalized to the quantum interval-valued expectation value. This
is used to prove an imprecise Kochen-Specker theorem \cite{BELL_1966,kochenspecker1967,Redhead1987-REDINA,peres1995quantum,Jaeger2007}
which suggests a possible resolution of the Meyer-Mermin debate on
the impact of finite-precision measurement on the Kochen-Specker theorem
\cite{PhysRevLett.83.3751,Mermin1999}.
\end{comment}
\begin{comment}
We present a mathematical framework based on quantum interval-valued
probability measures to study the effect of experimental imperfections
and finite precision measurements on defining aspects of quantum mechanics
such as contextuality and the Born rule. While foundational results
such as the Kochen-Specker and Gleason theorems are valid in the context
of infinite precision, they fail to hold in general in a world with
limited resources. Here we employ an interval-valued framework to
establish bounds on the validity of those theorems in realistic experimental
environments. In this way, not only can we quantify the idea of finite-precision
measurement within our theory, but we can also suggest a possible
resolution of the Meyer-Mermin debate on the impact of finite-precision
measurement on the Kochen-Specker theorem.
\end{comment}
Sec.~\ref{sec:Interval-Uncertainty} starts from reviewing how to
formulate classical ``imprecise probabilities'' \cite{Shafer1976,GilboaSchmeidler1994,Marinacci1999,JamisonLodwick2004,HuberRonchetti2009,Grabisch2016}
as interval-valued probability measures (IVPMs). Given a classical
IVPM, its core is the set of the real-valued probability measures
consistent with it \cite{Shapley1971,GilboaSchmeidler1994,NgMoYeh1997Chinese,Grabisch2016},
and the expectation value with respect to it is just classical Choquet
integral \cite{Choquet1954,GilboaSchmeidler1994,Grabisch2016}. These
classical properties can naturally be extended for QIVPMs when we
extend classical IVPMs to QIVPMs. In Sec.~\ref{sec:Kochen-Specker},
we utilize the previous defined expectation values while proving the
finite precision variants of the Kochen-Specker theorem \cite{BELL_1966,kochenspecker1967,Redhead1987-REDINA,peres1995quantum,Jaeger2007},
and establish bounds on its validity in realistic experimental environments.
In Sec.~\ref{sec:Gleason}, we are able to show that, while a QIVPM
incorporating the effects of finite precision might not be consistent
with Gleason's unique state~$\rho$ on all projectors defined on
a Hilbert space~$\Hilb$ of dimension $d\ge3$, there is a mathematically
precise sense in which one recovers the original Gleason theorem asymptotically.
Specifically, it is possible to construct a class of QIVPMs representing
bounded resources that is parameterized by the size of the intervals.
We then demonstrate that all QIVPMs in this class are consistent with
a non-empty “ball” of quantum states whose radius is defined
by the maximal length of the intervals characterizing the uncertainties.
As the size of the intervals goes to zero, this ball of quantum states
converges to a point representing the unique state consistent with
the Born rule and Gleason's theorem.

@@

\begin{tabularx}{\linewidth}{YYYYYYY}
\toprule 
\addlinespace
 & Conventional & Modal & Discrete (I) & Discrete (II) & QPMFF & QIVPM\tabularnewline\addlinespace
\addlinespace
\midrule
\midrule 
\addlinespace
 & Chapter~\ref{chap:Conventional-Quantum-Theory} & Sec.~\ref{modalquantum} & Sec.~\ref{discretequantumtheoryI} to \ref{sec:Geometry-of-Entangled-States} & Sec.~\ref{sec:Discrete-Quantum-Theory-(II)} & Sec.~\ref{sec:Toward-IVPM} & Chapter~\ref{chap:QIVPM}\tabularnewline\addlinespace
\addlinespace
\midrule
\addlinespace
States space & $\mathbb{C}^{D}$ & $\Fq^{D}$ & $\Fpp^{D}$ & Local region in $\Fpp^{D}$ & $\Fpp^{D}$ & $\mathbb{C}^{D}$\tabularnewline\addlinespace
\addlinespace
\midrule
\addlinespace
Likelihood of events is predicted by & Real-valued probability & Possible or impossible & Possible or impossible & Cardinal probability & Real-valued probability, but no sensible Born rule & Interval-valued probability\tabularnewline\addlinespace
\addlinespace
\midrule
\addlinespace
Expectation value & Defined & Undefined & Formally defined & Undefined &  & Defined\tabularnewline\addlinespace
\addlinespace
\midrule
\midrule 
\addlinespace
 &  & Sec.~\ref{modalquantumcomputing} & Sec.~\ref{discretequantumcomputingI} & Sec.~\ref{discretequantumcomputingII} &  & \tabularnewline\addlinespace
\addlinespace
\midrule
\addlinespace
Deutsch's algorithm & Yes & Maybe no & Yes & Yes &  & \tabularnewline\addlinespace
\addlinespace
\midrule
\addlinespace
Efficiently solve $\usat$ & Unlikely & Yes & Partially & Unlikely &  & \tabularnewline\addlinespace
\addlinespace
\midrule
\addlinespace
Grover search algorithm & Yes &  &  & Yes &  & \tabularnewline\addlinespace
\addlinespace
\bottomrule
\end{tabularx}
\end{sidewaystable}
The organization of this thesis is summarized in Table~\ref{tab:organization},
and the further extended questions are in final Chapter~\ref{chap:Further-Discussion}.

@@

\chapter{Further Discussion\label{chap:Further-Discussion}}

As we discussed in previous section, the finite-precision extension
of the Gleason theorem, Thm.~\ref{thm:Finite-precision-Gleason},
only applies on a class of QIVPMs inducing by composting interval
maps $f:\left[0,1\right]\rightarrow\mathscr{I}$ and quantum probability
measures $\muB_{\rho}:\events\rightarrow\left[0,1\right]$.

\begin{comment}
 Almost all QIVPMs in Chapter~\ref{chap:QIVPM} are closely related
to quantum probability measures. For example, given a $\delta$-deterministic
QIVPM, Lemma~\ref{lemma:delta2zero} provides a way to derive a $0$-deterministic
QIVPM $\barmuD:\events\rightarrow\left\{ \imposs,\necess\right\} $
which is essential a quantum probability measure $\mu:\events\rightarrow\left[0,1\right]$
satisfying $\mu\left(P\right)$ is either $0$ or $1$. Moreover,
an interval map $f:\left[0,1\right]\rightarrow\mathscr{I}$ in Sec.~\ref{sec:Gleason}
provides an systemic way to derive QIVPMs from quantum probability
measures.

the only QIVPM unrelated to quantum probability measures is defined
in Table~\ref{tab:non-Born-QIVPM}, and is ill-behaved as having
an empty core.

On one hand, classical IVPMs could be studied directly without deriving
from or to classical probability measures.

On the other hand, to state the original Gleason theorem, we don't
need to know a state a priori, and use this state to derive a quantum
probability measures and further a QIVPM. Therefore, it would be desirable
if we could answer the following variant of Gleason theorem which

whether shirking the maximum length of intervals alone could provide
a well-behaved QIVPMs. More precisely, we have the following question.

\begin{question}In a Hilbert space $\Hilb$ of dimension $D\geq3$,
given a QIVPM~$\bar{\mu}:\events\rightarrow\mathscr{I}$, $\alpha$
be the maximum length of intervals in $\mathscr{I}$, is there an
$\varepsilon\in\left(0,1\right)$ such that $\alpha<\varepsilon$
implies $\coreBorn\left(\bar{\mu},\events\right)\ne\emptyset$.\end{question}
\end{comment}
\begin{comment}

@@

\todo{Add a two-page conclusion chapter which could include further
discussion, but it can be only a few sentences? }\todo{Discrete
Schrödinger equation? And its relation with Grover's algorithm? }\todo{Notice
that in differential geometry, Killing–Hopf theorem asserts that having
positive constant curvature guarantees a manifold is essentially a
sphere, but say little things even if the curvature has only small
deviation from constant. Since a quantum probability measure is glued
by classical probability measures, like a manifold. When their relations
are exact, Gleason's theorem asserts that a quantum probability measure
can essentially induced by the Born rule, but they can be wild when
we move on to interval-valued situation! }
\end{comment}
